### 2024-06-16 thesis/ms-fletcher-porter.tex
Increasing the page count

%% Use this if you want to get links and nice output. Works well with pdflatex.
\usepackage{hyperref}
\hypersetup{
  pdfpagemode=UseNone,
  pdfstartview=FitH,
  colorlinks=true,
  urlcolor=xkcd_blue,
  linkcolor=black,
  citecolor=black,
  pdftitle=A Model for Selecting Software Tools for Fluid Power Systems,
  pdfauthor=Fletcher Porter,
  pdfkeywords={
    Mechatronics,Control Engineering,Software Tooling,Stewart Platform,Pneumatic actuator
  }
}

@@

\usepackage{siunitx}
\usepackage{gensymb}
\usepackage{svg}
\usepackage{lipsum}  % for placeholder text
\usepackage{subcaption}

@@

%% Your thesis title comes here and again before a possible abstract in
%% Finnish or Swedish . If the title is very long and latex does an
%% unsatisfactory job of breaking the lines, you will have to force a
%% linebreak with the \\ control character.
%% Do not hyphenate titles.
%%
\thesistitle{A Model for Selecting Software Tools for Fluid Power Systems}

@@

A research group at Okayama University of Science has, over the past years, developed a novel soft, pneumatic actuator that extends when supplied with pressure called an EFPA~\cite{Noritsugu2005}. This section will study the tech stacks for three applications of this actuator, a six-legged mobile robot for providing the elderly with an interactive physical activity~\cite{Hase2020}, a pipe-traversal robot for inspecting and repairing water pipe networks~\cite{Shinohara2020}, and a lightweight servo valve based on bending hoses to support EFPA applications~\cite{Kobayashi2020}. All of these can be seen in figure~\ref{fig:soft_pneumatic_actuators_demo}. Note that because this is a case study on work we did not participate in, some details must be inferred.

@@

\subsection{A Novel Stack for a Pneumatic Stewart Platform}

@@

Finally, the control layer had to use the API of the middleware layer to implement the business logic. One case of that logic was to have a controller to regulate the position of each cylinder. It had to be implemented in Rust to interface with the Ethercrab library. It also required some code to initialize the EtherCAT communication and to configure the Beckhoff terminals. There's also a user interface to interactively set the cylinder position. An interesting issue came up where the position sensor signal, mapping the $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ IO to the range 0..1~\verb|float|, but only reached the range 0..0.5~\verb|float|. This was because the step-down resistor was larger than designed since the right size wasn't available, so the maximum sensor reading was smaller than $\qty{10}{\volt}$. This unhandled gain is an undesired interface between the Control and IO layers since it demands the controller be written with knowledge of the minutia of how the IO was implemented. Further research will go toward how to prevent this.

The stack used for this pneumatic Stewart platform was Beckhoff EtherCAT Terminals with supplemental circuits for IO, a Raspberry Pi computer, Linux/Ethercrab middleware, and the control layer was built with Rust.

@@

The mechatronic tech stack theory basically held to scrutiny through the case studies.  All cases presented all layers of the stack except the EFPA robots, which had no middleware layer, though that is an interesting insight. With this model, the commissioning issues with the steering linkage Dolores could partly be diagnosed as too-tight coupling between layers.  Commissioning the Stewart platform was basically easy; by separating functions into different categories, work on any layer could be done independently.  It also raises compelling questions; can Dolores' Beckhoff industrial computer be replaced with a Raspberry Pi (only if it would support Beckhoff's middleware)?  Could the controller for the Stewart platform have been written in C or a PLC language?  What would it take to change the fieldbus protocol without the hassle seen with Dolores?  These questions are natural to ask when the tooling is presented in a structured manner as with this model.

The main issue that arose was the IO layer reaching into higher layers.  Dolores' old control system failed because it couldn't support a change in communication protocol.  The new system is over-complex because the control tools struggle to support all communication protocols.  The Stewart platform also saw issues with this; measurement signals were unexpectedly low due to an off-size resistor in the IO layer which impacted control function.  Certainly, some knowledge of the details of the IO layer's implementation must be communicated to the middleware or control layers such that correct middleware is run and sensor calibrations are accounted for.  How this could be implemented will be studied in future work.

An additional issue is that the computer layer isn't overly interesting.  Each of the case studies simply used whatever computers are available or easy to use.  In some circumstances, it could be more interesting, like to discuss which computer architecture to use for an embedded application, or whether to use an embedded Linux system like a Raspberry Pi or bare metal, or even a simple EEPROM.  Perhaps the computer layer could steal operating systems away from the middleware layer, or these case studies weren't interesting for that layer.

@@

This paper presented three case studies to demonstrate the utility of the novel theory of a mechatronic tech stack. These cases were diverse in their applications and methods and by viewing them through the lens of the tech stack, new knowledge of how fluid power systems can be commissioned was gained. Also, by studying several stacks, we have formed a pool of tools to draw from. The mechatronic tech stack can be re-rendered to reflect this, as here in figure~\ref{fig:mechatronic_tech_stack_filled}.

>>>>

%% Use this if you want to get links and nice output. Works well with pdflatex.
\usepackage{hyperref}
\hypersetup{
  pdfpagemode=UseNone,
  pdfstartview=FitH,
  colorlinks=true,
  urlcolor=xkcd_blue,
  linkcolor=black,
  citecolor=black,
  pdftitle=A Model for Selecting Software Tools for Mechatronic Systems,
  pdfauthor=Fletcher Porter,
  pdfkeywords={
    Mechatronics,Control Engineering,Software Tooling,Stewart Platform,Pneumatic actuator
  }
}

@@

\usepackage{siunitx}
\usepackage{gensymb}
\usepackage{svg}
\usepackage{lipsum}  % for placeholder text
\usepackage{subcaption}
\usepackage{bigfoot} % for \verb in \footnote

@@

%% Your thesis title comes here and again before a possible abstract in
%% Finnish or Swedish . If the title is very long and latex does an
%% unsatisfactory job of breaking the lines, you will have to force a
%% linebreak with the \\ control character.
%% Do not hyphenate titles.
%%
\thesistitle{A Model for Selecting Software Tools for Mechatronic Systems}

@@

A research group at Okayama University of Science has, over the past years, developed a novel soft, pneumatic actuator that extends when supplied with pressure called an EFPA~\cite{Noritsugu2005}. This section will study the tech stacks for three applications of this actuator, a six-legged mobile robot for providing the elderly with an interactive physical activity~\cite{Hase2020}, a pipe-traversal robot for inspecting and repairing water pipe networks~\cite{Shinohara2020}, and a lightweight servo valve based on bending hoses to support EFPA applications~\cite{Kobayashi2020}. All of these can be seen in figure~\ref{fig:soft_pneumatic_actuators_demo}. Note that because this is a case study on work I did not participate in, some details must be inferred.

@@

\subsection{A Novel Stack for a Pneumatic Stewart Platform}\label{sec:stewart_platform}

@@

Finally, the control layer had to use the API of the middleware layer to implement the business logic. One case of that logic was to have a controller to regulate the position of each cylinder. It had to be implemented in Rust to interface with the Ethercrab library. It also required some code to initialize the EtherCAT communication and to configure the Beckhoff terminals. There's also a user interface to interactively set the cylinder position. An interesting issue came up where the position sensor signal, mapping the $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ IO signal to the range 0..1~\verb|float|, but only reached the range 0..0.5~\verb|float|. This was because the step-down resistor was larger than designed since the right size wasn't available, so the maximum sensor reading was smaller than $\qty{10}{\volt}$. This unhandled gain is an undesired interface between the Control and IO layers since it demands the controller be written with knowledge of the minutia of how the IO was implemented. Further research will go toward how to prevent this.

Also interesting in the architecture of a control system is what development environments can be constructed for it.  The development environment is distinct from the heretofore discussed mech stack since it is not a part of the function of the running system, but instead is means which an engineer uses to turn an idea into assets for a mechatronic system.  For a given tech stack, any development environment generally can in principle yield the same assets, but a good tech stack may lend itself to higher quality, more consistent output, faster development cycles, or at least to be enjoyable for individual engineers to use.

For the control layer, I wrote the code in the Sublime Text text editor~\cite{SublimeText} and configured it to automatically format the source code according to standards for the Rust language using the \verb|rustfmt| utility~\cite{rustfmt}.  Other text editors like Emacs, Vim, or VSCodium would have also been good choices, but Sublime is what is easiest for me.  This was done on a Windows workstation, but it was compiled and run on the Linux control computer.  To achieve this, I checked the changes on the workstation into a \verb|git| version control repository and pushed them to the control computer\footnote{A small issue that comes up is that \verb|git| generally doesn't allow users to push changes to a branch of a repository that the local user has checked out, which is the case here~\cite{GitConfigReceiveDenyCurrentBranch}.  This is to prevent data loss.  To enable this, \verb|git| can be configured to update the local branch only if there is no possibility of data loss by executing \verb|git config receive.denyCurrentBranch updateInstead|.}.  Then, though a remote terminal, I executed commands to compile and run the program.  This became quite annoying with how many steps there are, so I use the utility \verb|cargo-watch| to automatically execute the compile and run commands on the control computer when there are updates\footnote{The command was similar to \verb|cargo watch -x 'run -- eth0'|.}~\cite{CargoWatch}.  This still is somewhat tedious and so requires more development.  It also would have been mitigated by developing in some kind of Linux environment or by using a containerized environment like would be provided by Docker or Vagrant, which would have allowed running some tests on the development machine~\cite{DockerDev,VagrantDev}.  Most of the middleware was prepared with the Raspberry Pi OS image that was installed while comissioning the control computer, but there was some additional configuration, e.g. to install \verb|git|, that was done with the OS's \verb|apt| utility~\cite{apt}.  As well, library depndencies for the control program like Ethercrab were managed with the Rust \verb|cargo| utility~\cite{Cargo}.  The computer, once commissioned, did not require any development.  The IO layer's development was aided by a four-channel oscilliscope and a multimeter to validate electrical connections and ensure signal were propogating as expected.  It was also interesting to observe the timing jitter of the program, which was generally very small, on the order of microseconds.

The in-depth consideration of the development environment was possible because of the open nature of all the tools considered.  Every tool, both a part of the control system and the development environment, have interfaces that enable and encourage users to compose or develop tools to make an environment that meets their needs.  This same discussion could not be had with regards to Dolores since its PLC-based system was very closed off, and any interfacing demanded mastering of low-level communicaiton protocols.  The development environment was simply whatever Beckhoff blessed; no more, no less.

This Stewart platform employed a tech stack consisting of Beckhoff EtherCAT terminals with supplementing circuits for IO, a Raspberry Pi computer, Linux and Ethercrab for middleware, and control layer built with the Rust programming language.  By crafting a custom stack with largely open components, it was easy to also consider and iterate on the development process.  This all proved to be quite successful, as all the development, including ideating on the tech stack itself and learning how to use Ethercrab, was completed in only a week and a half.

\subsection{Stress Testing IO Requirements}

@@

This paper has introduced a theory of a mechatronic tech stack which describes function and tooling for mechatronic systems, namely as discrete layers of control, middleware, computer, and IO.  This theory is motivated by analogy to the web stack, used in IT software development, and to the OSI networking model, used to describe networked computer systems.  The OSI model also provided a means of discriminating the layers that make up the model, importantly by finding contrasts in the morphology, syntax, and semantics of the behavior of components of the control system.

It also tested this theory against several case studies.  In one case, a hydraulic steering linkage testbed showed that a change in one absraction layer can demand rearchitecture of the other abstraction layers if the control system is not carefully designed.  In another, an organization employed variations on one tech stack to build robots based on soft actuators, presumably to capitalize on institutional knowledge of that stack.  The last used the development of a stewart platform to show how a control system can be developed based on this theory.  By organizing the tools used in these case studies based on the function they perform and the interface they provide, I generated these compelling insights into the architecture of the control system.

Although useful, the theory presented is not yet mature.  The principal issue is that the definitions of the abstraction layers and the boundaries between them lack specificity.  For instance, I state that the computer layer is the hardware that performs computation and that the interface it provides is its architecture.  What constitutes such computational hardware, and what exactly is a computer architecture?  And what is a practitioner meant to do with this information?  Stating that a computer uses a SuperH or x86 architecture doesn't immediately suggest which is better for the application, and likewise for a ATmega328P or ATmega32u4 computer chip.  Providing specific definitions for the layers and their boundaries would be useful.

To increase the specificity of the definition of they theory, motivating its definition with more background theory could be appealing.  Presently, it's motivated by an analogy to the web stack and by a test defined in the OSI model.  If even more theories were employed to discern the structure of the stack and they pointed to similar results, that would substantially strengthen the argument for this theory.  A concrete step towards a specific definition for the tech stack theory would be to emulate the layer definitions provided in the OSI model, which enumerate the purpose, provided services, and function of each abstraction layer~\cite[§7]{ISO7498-1}.  However, it would be better to draw from more sources than this one standard for how to structure the definitions in this theory.  Further benefit could be found by studying and creating more case studies to challenge the structure of the theory and to test its utility.

This theory can also be extended.  Throughout this work, it has been called a \textit{mechatronic} tech stack, but truly it's a control system tech stack.  To make it a true mechatronic tech stack, it should include tooling for mechanical and electronic systems.  A mechanical section of the stack would likely grow from the bottom of the stack shown in figure~\ref{fig:mechatronic_tech_stack}, not colliding with the current stack, but an electrical component would likely be more complicated, since it is needed by both the mechanical stack for transducers and sensors, and by the control stack since the computer \textit{is} an electrical component, and the IO layer often needs some elecrical finesse, as noted in §\ref{sec:stewart_platform}.  It could also be extended in the software realm; the production of graphical monitoring systems and tele-operation programs generally require substantial non-control software which could be described by different layers, perhaps replacing the control layer in the present stack.

This theory would also benefit from the study of adjacent work that doesn't develop the mechatronic tech stack itself.  This theory describes the function of mechatronic tooling, but it provides no notion of what tools are best for a particular application.  In evaluating the case studies, some initial guesses were developed, like that tools should encourage loose coupling, take advantage of institutional knowledge, and be easy to troubleshoot.  This can be formalized by interviewing practitioners and querying what qualities they value in tools as well as by studying extant literature on the subject of tool quality.

As well as considering the quality of a tool, it's important also to study the legal existance of mechatronic tools.  Almost always when making and redistributing mechatronic systems and related design assets, engineers and their associated organizations must agree to some license agreement with the authors of their tools, with these legal agreements being underpinned by various intelectual property laws.  While it is tempting to just ``hit agree'', these license agreements can place notable conditions on the redistribution of mechatronic systems.  A notable subset of these agreements are those \textit{open-source} licenses, which are often discussed as interesting to practitioners in mechatronics, but the legal requirements for open-source software are often ignored or misunderstood.  Open-hardware licenses are even less understood.  A deep dive into law around mechatronic tools would greatly enhance the mechatronic tech stack theory by allowing practitioners to evaluate if they're even allowed to use a tools, regardless of its quality.

Finally, this theory will not do much benefit locked away in a masters thesis that so few people will ever read.  For it to be truly useful, it must be deployed widely among practitioners of mechatronics.  This can be accomplished in several ways.  The theory and progress on it can be published to journals and conferences (indeed, a subset of this thesis is in review for the Japanese Fluid Power Society's conference~\cite{Porter2024}) or perhaps even to more informal forums like a personal blog or social media.  Collaborating on projects with other research groups and industry and deploying this theory will both spread the ideas of the thesis and test its utility.  Finally, developing teaching for this project would be an useful test of its utility.  Students who don't have much experience with architecting mechatronic systems will have the most useful feedback to whether the structure provided by this theory is truly useful for developing projects.

@@

This paper presented three case studies to demonstrate the utility of the novel theory of a mechatronic tech stack. These cases were diverse in their applications and methods and by viewing them through the lens of the tech stack, new knowledge of how fluid power systems can be commissioned was gained. Also, by studying several stacks, I have formed a pool of tools to draw from. The mechatronic tech stack can be re-rendered to reflect this, as here in figure~\ref{fig:mechatronic_tech_stack_filled}.
